hbm_paper_results

### 2024-01-06 paper_main_results.Rmd
splitting out big rmd file

@@

```{r}
subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
  geom_point(alpha=0.3) +
  geom_smooth() +
  facet_grid(dataset ~ sim.site, scales="free_y")
```

@@

```{r}
#apply
perm_t.tests <- mclapply(diffs_subj.list, centile.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_subj.list))
names(perm_t.tests) <- n_perm_list
cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
fwrite(cent_t.tests.p, file="derivatives/ukb_perm_csvs/centile_t.tests.csv")

centile.t.tests(perm_subj_list[[1]])
```

@@

Little plots
```{r}
#define plotting
plot.cent.diffs <- function(df){
mean.diffs.long <- df %>%
  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

cent.diffs.plot <- mean.diffs.long %>%
  dplyr::filter(dataset != "raw") %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 11) +
  facet_wrap(~pheno_cat) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
  ggtitle("Figure 2. Absolute errors in centile scores across ComBat processing pipelines")

```{r}
lapply(perm_diffs, plot.cent.diffs)
```

W/in subj average plots
```{r}
mean.subj.centile.err <- function(df){
mean.diffs.subj <- df %>%
  mutate(dataset = factor(dataset, levels = c("cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
  rowwise() %>%
  dplyr::mutate(mean = mean(c_across(starts_with("abs.diff"))))

return(mean.diffs.subj)
#nrow(mean.diffs.subj) == (n_unique(mean.diffs.subj$participant) * 4) #check length
}
perm_cent.err <- mclapply(perm_diffs, mean.subj.centile.err, mc.preschedule = FALSE)
names(perm_cent.err) <- n_perm_list
cent_err.p <- bind_rows(perm_cent.err, .id = "permutation")

#write out for plotting later
saveRDS(cent_err.p, "ukb_perm_csvs/centile_errors.RDS")
```

```{r}
cent.diffs.v.plot <- cent_err.p %>%
  mutate(mean = mean*100) %>% #for visuals
  ggplot() +
  geom_violin(aes(x=dataset, y=mean, fill=permutation)) +
  labs(y= "Magnitude Centile Error", x = "ComBat Configurations") +
  ggtitle("Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
```

>>>>

plot w/in feature average centile errors

```{r}
#define plotting
plot.cent.diffs <- function(df){
mean.diffs.long <- df %>%
  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  group_by(dataset) %>%
  dplyr::summarize(across(starts_with("abs.diff_"), mean)) %>%
  ungroup() %>%
  dplyr::select(-ends_with(".z")) %>% #drop z-score cols
  pivot_longer(cols=starts_with("abs.diff_"), values_to="centile_abs.diff", names_to = "pheno", names_prefix = "abs.diff_") %>%
  mutate(pheno_cat = as.factor(case_when(
      pheno %in% vol_list_global ~ "Global Volume",
      pheno %in% vol_list_regions ~ "Regional Volume",
      pheno %in% sa_list ~ "Regional SA",
      pheno %in% ct_list ~ "Regional CT",
      TRUE ~ NA_character_)),
      label = sub("_[^_]*_", "_", pheno))

cent.diffs.plot <- mean.diffs.long %>%
  dplyr::filter(dataset != "raw") %>%
  mutate(centile_abs.diff = centile_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_histogram(aes(x=centile_abs.diff, fill=dataset, y = stat(width*density)), color="black", alpha=0.6, position="identity") +
  theme_linedraw(base_size = 11) +
  facet_wrap(~pheno_cat) +
  scale_y_continuous(labels = scales::percent) + 
  scale_fill_manual(values=dataset_colors, name = "ComBat \nConfiguration", labels=dataset_names) +
  labs(y= "Percent Features", x = "Mean Absolute Centile Error") +
  ggtitle("Absolute errors in centile scores across ComBat processing pipelines")

@@

```{r}
subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
  geom_point(alpha=0.3) +
  geom_smooth() +
  facet_grid(~dataset, scales="free_y")

subj_sum.df %>%
  dplyr::filter(dataset == "cf.gamlss" | dataset == "cf.gam" | dataset == "cf.lm") %>%
  ggplot(aes(x=mean_centile, y=mean_cent_diff, color=sex)) +
  geom_point(alpha=0.3) +
  geom_smooth() +
  facet_grid(dataset ~ sim.site, scales="free_y")
```

test to see if slope of non-dominant group is nonzero!!!

@@

```{r}
#apply
perm_t.tests <- mclapply(diffs_subj.list, centile.t.tests, mc.preschedule = FALSE, comp_multiplier=length(diffs_subj.list))
names(perm_t.tests) <- n_perm_list
cent_t.tests.p <- bind_rows(perm_t.tests, .id = "column_label")
fwrite(cent_t.tests.p, file="derivatives/ukb_perm/centile_t.tests.csv")
```

@@

Little plots

```{r}
lapply(diffs_subj.list, plot.cent.diffs)
```

```{r}
dataset_names <- c(ukb_CN = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
cent.diffs.v.plot <- perm_subj.df %>%
  mutate(mean_cent_abs.diff = mean_cent_abs.diff*100) %>% #for visuals
  ggplot() +
  geom_violin(aes(x=dataset, y=mean_cent_abs.diff, fill=perm), scale="count") +
  labs(y= "Mean Magnitude Centile Error", x = "ComBat Configurations") +
  scale_x_discrete(labels=c(dataset_names)) +
  guides(fill=guide_legend(title="Sampling \nPermutation")) + 
  ggtitle("Fig 1B. Within-subject mean magnitude of error in centile scores across ComBat \nconfigurations replicated 10x")
cent.diffs.v.plot
ggsave(filename="figs/centile_violin.png", cent.diffs.v.plot)
```
